---
title: "1-CalculateInhibitionScore"
output: html_document
date: "2024-10-15"
editor_options: 
  chunk_output_type: console
---
```{r, echo=FALSE}
setwd("/Users/Julian/Library/CloudStorage/OneDrive-SmithsonianInstitution/Spotted Salamander 2024/#R-files")
setwd("C:/Users/Owner/OneDrive - Smithsonian Institution/Spotted Salamander 2024/#R-files")
Bd <- read.csv("LongformAllPlates.csv")
```


### Remove wells
LOOK at each plate and make sure no contaminates OR extreme outliers within a sample type remove those here so there are notes of which wells were removed. Remove wells with no treatment (ie. no mucosome only Bd). WSBP-17 (Plate 2 2C) was the only well with visible contaminants during the growth assay, all others were not included in the assays if we didn't have enough mucosome sample leftover. NDPC plate 2 11A was excluded as an outlier (> 0.500 OD)
```{r, warning=FALSE, message=FALSE}
library(dplyr)
Bd <- Bd %>% filter(!(AmphibianID %in% c('AZU-A-11', 'AZU-13',
                                         'NAT-A-04', 'NAT-A-03', 'ORI-17')),
                    !(Plate == 2 & Column == 2 & Row == 'C'),
                    !(Plate == 2 & Column == 10 & Row == c('C', 'D','G','H')),
                    !(Plate == 2 & Column == 6 & Row == c('D','G','H')),
                    !(Plate == 2 & Column == 7 & Row == c('D','H')),
                    !(Plate == 2 & Column == 8 & Row == 'D'),
                    !(Plate == 2 & Column == 9 & Row == 'H'),
                    !(Plate == 2 & Column == 11 & Row == 'A'),
                    !(Plate == 3 & Column == 10 & Row == c('E','F','G','H')),
                    !(Plate == 3 & Column == 1 & Row == c('D','E','F','G','H')),
                    !(Plate == 3 & Column == 1 & Row == c('C','D','G','H'))
                    )
```

For each plate, For each row in the isol df 
```{r}
for(i in 1:nrow(Bd)){
# Take the OD of the row i and divide by the AveND -1
# assign it to Inhibit column in isol df
  Bd$T_OD[i] <-  log((Bd$OD[i] / (1-Bd$OD[i])+1)) 
  
}
```
  
### Bd growth plots
Level off of positive growth seen ~ day 7, we will exclude days 11 and 12 from analysis
```{r, warning=FALSE}
# histogram
hist(Bd$OD, breaks = 30)

# This transformation makes the data more linear, better to fit linear model
hist(Bd$T_OD, breaks = 30)

# Look at positive control growth
PC <- Bd[Bd$AmphibianID == c("NDPC", "PC"),]

library(ggplot2)

plot12 <- ggplot(PC, aes(Day, T_OD)) +
  geom_point(aes(color = AmphibianID)) +
  geom_smooth(method = loess)

plot12 + facet_grid(~Plate~AmphibianID, scales = "free")
```

We want to calculate the slope of growth for each well in our assays. We want to take an average of those slopes for each peptide concentration or controls.

Then we want to divide the average slope of each sample-Bd combo by the nutrient-depleted positive control for the specific Bd isolate and then subtract one and we get our inhibition score!!!!!!

Carly's old code she wrote for this also subset within Bd isolates (Muletz-Wolz 2017, Frontiers). Just keeping that code as is, but we only have 1 isolate.

### Manipulate dataset for linear regression
```{r}
# First let's merge column and row to make a well column
Well <- as.factor(paste(Bd$Row, Bd$Column, sep = ""))
Bd <- cbind(Bd,Well)

# Subset to days based on PC growth
Bd <- subset(Bd, Day %in% c(0,1,4,5,7))

# Make a column to put in calculated slopes
Bd$Inhibit_slope <- NA

# specify unique plates
plate_unique <- unique(Bd$Plate)

# specify our intercept as 0
intercept = 0  
```

### Linear model loop
```{r}
# For each plate
for(i in 1:length(plate_unique)){
  Plate_subset <- subset(Bd, Plate == plate_unique[i]) 
  iso <- unique(Plate_subset$Isolate)
  
  # For each of unique isolates 
  for(j in 1:length(iso)){
    # Subset to only that iso i and assign it to a new dataframe isol
    isol <- subset(Plate_subset, Isolate==iso[j])
    well <- unique(isol$Well)
    
    # For each unique well
    for(k in 1:length(well)){
      # Subset to that well and assign it to a new dataframe well_subset
      well_subset <- subset(isol, Well == well[k])
      # Make a linear model for each bacteria based on the transformed OD
      lm_well <- lm(I(well_subset$T_OD - intercept) ~ 0 + well_subset$Day)
      ## then take the slope and put into appropriate column in isol df
      ## if we didn't force the intercept slope is in [2]
      ## well_subset$Inhibit_slope <- lm_well$coefficients[2]
      well_subset$Inhibit_slope <- lm_well$coefficients[1]
      well_subset$R_squared <- summary(lm_well)$adj.r.squared
      well_subset$R_squared <- ifelse(well_subset$R_squared < 0.2, NA,
                                      well_subset$R_squared)
      well_subset$Inhibit_slope <- ifelse(well_subset$R_squared < 0.2, NA,
                                          well_subset$Inhibit_slope) 
      # Once that is all completed for that unique isolate, assign it into the original df 
      isol[isol$Well==well[k],"Inhibit_slope"] <- well_subset$Inhibit_slope
      isol[isol$Well==well[k],"R_squared"] <- well_subset$R_squared
      ## get the average of ND inhibit_slope
      aveND <- mean(isol$Inhibit_slope[isol$AmphibianID == "NDPC"])
      
      ## for each row in the isol df
      for(l in 1:nrow(isol)){
        # Take the inhibit_Slope of the row l and divide by the Avg ND slope and then -1, 
        # Assign to new column called final_inhibition
        isol$final_inhibition[l] <- 1- (isol$Inhibit_slope[l]/aveND)
      }
    }
    # Once that is all completed for that unique isolate, assign it into the original df 
    Plate_subset[Plate_subset$Isolate==iso[j],"Inhibit_slope"] <- isol$Inhibit_slope
    Plate_subset[Plate_subset$Isolate==iso[j],"final_inhibition"] <- isol$final_inhibition
    Plate_subset[Plate_subset$Isolate==iso[j],"R_squared"] <- isol$R_squared
  }
  Bd[Bd$Plate==plate_unique[i],"Inhibit_slope"] <- Plate_subset$Inhibit_slope
  Bd[Bd$Plate==plate_unique[i],"final_inhibition"] <- Plate_subset$final_inhibition
  Bd[Bd$Plate==plate_unique[i],"R_squared"] <- Plate_subset$R_squared
}
```

### Write out data .csv
we have replication of the final_inhibition for every day, need to subset only one day
```{r}
write.csv(Bd, "InhibitionScoresAllMucosome.csv")

Bd_scores_2 <- Bd %>% 
  select(-X, -X.1, -X.2) %>% 
  filter(Day == 7) %>% 
  mutate(final_inhibition_percent = final_inhibition * 100)

write.csv(Bd_scores_2, "InhibitionScoresMucosome.csv")
```

### Check your work!
```{r}
# subset to one plate and one well
Plate_subset <- subset(Bd, Plate == 1) 
well_subsetA4 <- subset(Plate_subset, Well == 'A4')

lm_wellA4 <- lm(I(well_subsetA4$T_OD - intercept) ~ 0 + well_subsetA4$Day)
summary(lm_wellA4)
plot((well_subsetA4$T_OD - intercept) ~ 0 + well_subsetA4$Day)

## Inhibition slope is the coefficient
lm_wellA4$coefficients[1]
Inhibit_slopeA4 <- lm_wellA4$coefficients[1]
summary(lm_wellA4)$adj.r.squared
aveND <- mean(Plate_subset$Inhibit_slope[Plate_subset$AmphibianID == "NDPC"])
1  - (Inhibit_slopeA4/aveND)

## plate 1 well A4 should have inhibit slope as 0.0338, inhibition score as -1.777
## and adjusted r-squared = 0.966
well_subsetA4
## We did it!

## Let's do one more random well 
Plate_subset <- subset(Bd, Plate == 3) 
well_subsetA4 <- subset(Plate_subset, Well == 'A4')

lm_wellA4 <- lm(I(well_subsetA4$T_OD - intercept) ~ 0 + well_subsetA4$Day)
summary(lm_wellA4)
plot((well_subsetA4$T_OD - intercept) ~ 0 + well_subsetA4$Day)

## Inhibition slope is the coefficient
lm_wellA4$coefficients[1]
Inhibit_slopeA4 <- lm_wellA4$coefficients[1]
summary(lm_wellA4)$adj.r.squared
aveND <- mean(Plate_subset$Inhibit_slope[Plate_subset$AmphibianID == "NDPC"])
1- (Inhibit_slopeA4/aveND)

## plate 4 well A4 should have inhibit slope as 0.0064, inhibition score as 0.436 
## and r-squared as 0.9258
well_subsetA4
## Confirmed!
```

